An Algorithmic Introduction to Lagrangian Coherent Structures

### 2014-01-31 main.tex
Correct some typos

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage[uniquename=init,firstinits,backend=bibtex,sorting=none]{biblatex}
\usepackage{calc}
\usepackage{color}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{listings}
\usepackage[numbered,framed]{mcode}
\usepackage[space]{grffile}
%% hyperref manual advises loading this package last
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,breaklinks=true}

@@

@@

The second step is to integrate the lambda-lines to obtain the Poincare return map. Integration of lambda-lines is performed with the $\eta_+$ and $\eta_-$ vector fields defined on the main grid. These vector fields are derived from eigenvalues and eigenvectors as described in~\textcite{farazmand12:_comput_lagran}. The eigenvector fields have orientation discontinuities, therefore the vector field must be locally reoriented for the integration. The algorithm used to perform variable step-size integration in vector fields with orientation discontinuities is enumerated in Table~\ref{t:variable step integration} and illustrated in Figure~\ref{f:variable step integration}. Linear interpolation is used when interpolating $\boldsymbol \eta_\pm$ within a grid element, since using higher order interpolation would necessitate verifying that there are no orientation discontinuities at more than only the four nearest grid points. Detection of orientation discontinuities is achieved by checking the inner product of the vector fields at adjacent grid points. Then, rotations exceeding 90\degree\, are classified as orientation discontinuities and are corrected before linear interpolation. Thus when setting the Cauchy-Green strain tensor main grid resolution, a histogram of eigenvector field rotations may be produced to ensure that all rotations are below 90\degree\, or close to 180\degree.

@@

Figures~\ref{f:ocean dataset hyperbolic shear lcs details strainline} and \ref{f:ocean dataset hyperbolic shear lcs details stretchline} show elliptic and hyperbolic LCS on 22 November 2006, i.e., the same time instant shown in~\textcite{haller13:_coher_lagran,beron-vera13:_objec_agulh}. Note that here the integration time is $T=30\,\mathrm{days}$, as opposed to $90$\,days in their analysis. Since for $90$\,days the hyperbolic LCS would be a confusing entanglement of lines, we choose a shorter integration time here, to be able to show hyperbolic LCS along with elliptic LCS. 
For the shorter integration time, the analysis with LCS Tool reveals a large coherent eddy bounded by a closed lambda-line at $(3,-32)$ which corresponds to eddy \#2 in Fig.\,3 of \textcite{beron-vera13:_objec_agulh}. Additionally, a further smaller coherent eddy core at $\sim(1,-31)$ is found that does not stay coherent over a time of $90$\, days. Strainlines (red) and stretchlines (blue) mark the core lines of strong normal attraction and repulsion.

>>>>

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage[uniquename=init,firstinits,backend=bibtex,sorting=none]{biblatex}
\usepackage{calc}
\usepackage{color}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{listings}
\usepackage[numbered,framed]{mcode}
\usepackage[space]{grffile}
\usepackage{lineno}
%% hyperref manual advises loading this package last
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,breaklinks=true}

@@

\linenumbers

@@

The second step is to integrate the lambda-lines to obtain the Poincare return map. Integration of lambda-lines is performed with the $\eta_+$ and $\eta_-$ vector fields defined on the main grid. 
%These vector fields are derived from eigenvalues and eigenvectors as described in~\textcite{farazmand12:_comput_lagran}.
0The eigenvector fields have orientation discontinuities, therefore the vector field must be locally reoriented for the integration. The algorithm used to perform variable step-size integration in vector fields with orientation discontinuities is enumerated in Table~\ref{t:variable step integration} and illustrated in Figure~\ref{f:variable step integration}. Linear interpolation is used when interpolating $\boldsymbol \eta_\pm$ within a grid element, since using higher order interpolation would necessitate verifying that there are no orientation discontinuities at more than only the four nearest grid points. Detection of orientation discontinuities is achieved by checking the inner product of the vector fields at adjacent grid points. Then, rotations exceeding 90\degree\, are classified as orientation discontinuities and are corrected before linear interpolation. Thus when setting the Cauchy-Green strain tensor main grid resolution, a histogram of eigenvector field rotations may be produced to ensure that all rotations are below 90\degree\, or close to 180\degree.

@@

Figures~\ref{f:ocean dataset hyperbolic shear lcs details strainline} and \ref{f:ocean dataset hyperbolic shear lcs details stretchline} show elliptic and hyperbolic LCS on 22 November 2006, i.e., the same time instant shown in~\textcite{haller13:_coher_lagran,beron-vera13:_objec_agulh}. Note that here the integration time is $T=30\,\mathrm{days}$, as opposed to $90$\,days in their analysis. Since for $90$\,days the hyperbolic LCS would be a confusing entanglement of lines, we choose a shorter integration time here, to be able to show hyperbolic LCS along with elliptic LCS. 
For the shorter integration time, the analysis with LCS Tool reveals a large coherent eddy bounded by a closed lambda-line at $(3,-32)$ which corresponds to eddy \#2 in Fig.\,3 of \textcite{beron-vera13:_objec_agulh}. Additionally, a further smaller coherent eddy core at $\sim(1,-31)$ is found that does not stay coherent over a time of $90$\, days. Both closed orbits are stable if the resolution for the Cauchy-Green strain tensor field is increased. Strainlines (red) and stretchlines (blue) mark the core lines of strong normal attraction and repulsion.
